---
title: "Installing additional solvers"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installing additional solvers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The **holiglm** package uses by default the **ECOS** solver.
However, utilizing commercial solvers such as

- [**MOSEK**](https://www.mosek.com),
- [**GUROBI**](https://www.gurobi.com),
- [**CPLEX**](https://www.ibm.com/products/ilog-cplex-optimization-studio/cplex-optimizer)

can significantly reduce the computation time.
Although these solvers are commercial, academic licenses are available free of charge for students and academic staff.
It is important to note that the source code for these solvers is not accessible from **CRAN** and must be obtained from the respective company's homepage.

Internally **holiglm** relies on the [**ROI**](https://cran.r-project.org/package=ROI) package, which provides a consistent interface across various solvers.
To use the commercial solvers, the corresponding **R** interfaces need to be installed.

## Installing MOSEK
To use the **MOSEK** solver in **hglm**, you need to install the **Rmosek** and **ROI.plugin.mosek** packages.
As **ROI.plugin.mosek** imports **Rmosek**, it is recommend to install **Rmosek** first.


### Installing **Rmosek**
Since **Rmosek** is commercial software, it must be obtained from [`https://www.mosek.com/`](https://www.mosek.com/).
Information about installing **Rmosek** on *Windows*, *MacOS* and *Linux* can be found at [`https://docs.mosek.com/latest/rmosek/install-interface.html`](https://docs.mosek.com/latest/rmosek/install-interface.html).

### Installing **ROI.plugin.mosek**
To install **ROI.plugin.mosek**, you need to have the **Rmosek** package installed.
Before installing **ROI.plugin.mosek**, make sure that **Rmosek** is correctly installed by loading it using the following command in **R**:
```{r, eval = FALSE}
library("Rmosek")
```
Once **Rmosek** is installed, you can install **ROI.plugin.mosek** directly from Gitlab using the following code:
```{r, eval = FALSE}
remotes::install_gitlab("roigrp/solver/ROI.plugin.mosek", INSTALL_opts = "--no-multiarch")
```

Since **ROI.plugin.mosek** only uses **R** code and does not link to any external library it is very unlikely that the installation will fail.
If you encounter difficulties during the installation,

1. Check if **Rmosek** was installed correctly, by executing `library("Rmosek")` in **R**.
2. Check your internet connection.


## Installing GUROBI
As the **gurobi** package is commercial software, it must be obtained from [`https://www.gurobi.com`](https://www.gurobi.com/).
The site [https://www.gurobi.com/documentation](https://www.gurobi.com/documentation/) provides installation instructions for **GUROBI** on *Windows*, *MacOS* and *Linux*.

### Installing **ROI.plugin.gurobi**
To install **ROI.plugin.gurobi**, you need to have the **gurobi** package installed.
Before installing **ROI.plugin.gurobi**, ensure that **gurobi** is correctly installed by loading it in **R**:
```{r, eval = FALSE}
library("gurobi")
```
Once **gurobi** is installed, you can install **ROI.plugin.gurobi** directly from Gitlab using the following code:
```{r, eval = FALSE}
remotes:::install_gitlab("roigrp/solver/ROI.plugin.gurobi", INSTALL_opts = "--no-multiarch")
```

Since **ROI.plugin.gurobi** only uses **R** code and does not link to any external library it is very unlikely that the installation will fail.
If you encounter difficulties during the installation,

1. Check if **gurobi** was installed correctly, by executing `library("gurobi")` in **R**.
2. Check your internet connection.


## Installing CPLEX
To use **ROI.plugin.cplex**:

1. Install [IBM ILOG CPLEX Optimizer](https://www.ibm.com/products/ilog-cplex-optimization-studio/cplex-optimizer)
2. Install [Rcplex](https://cran.r-project.org/package=Rcplex) from **CRAN**.
3. Install **ROI.plugin.cplex**

### Installing **ROI.plugin.cplex**
To install **ROI.plugin.cplex**, you need to have the **Rcplex** package installed. 
Before installing **ROI.plugin.cplex**, ensure that **Rcplex** is correctly installed by loading it in **R**:
```{r, eval = FALSE}
library("Rcplex")
```
Once **Rcplex** is installed, you can install **ROI.plugin.cplex** directly from Gitlab using the following code.
```{r, eval = FALSE}
remotes:::install_gitlab("roigrp/solver/ROI.plugin.cplex", INSTALL_opts = "--no-multiarch")
```

Since **ROI.plugin.cplex** only uses **R** code and does not link to any external library it is very unlikely that the installation will fail.
If you encounter difficulties during the installation,

1. Check if **Rcplex** was installed correctly, by executing `library("Rcplex")` in **R**.
2. Check your internet connection.
